---
title: "TCA metric description"
header-includes:
   - \usepackage{tikz}
    - \usetikzlibrary{shapes.geometric,arrows}
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

We first start by treating vehicles as point objects, P and Q. For the purpose of this paper, we have determined anti-clockwise rotations to be positive and clockwise rotations to be negative.  We define $\vec{u}$ and $\vec{v}$ as the velocities of P and Q respectively:

- $\vec{u} = u_0 \cos(\theta)\hat{i} + u_0\sin(\theta)\hat{j}$
- $\vec{v}=-v_0\sin(\phi)\hat{i}+v_0\cos(\phi)\hat{j}$

where:

- $\hat{i}$ and $\hat{j}$ are the axes of the vector plane
- $\theta$ is the angle of rotation of $\vec{u}$ at point P from $\hat{i}$
- $\phi$ is the angle of rotation of $\vec{v}$ at point Q from $\hat{j}$

To calculate TCA of points P and Q, we find the displacement from P to Q as a function of time. Then, in order to get the time at which the displacement is minimum, we differentiate the function, equate it to 0 and solve for time. This time is the time to closest approach (TCA) of P and Q. If P and Q collide with each other, the displacement at TCA will be 0.
To calculate P and Q at time t, we use the following notation:

\begin{equation}
    P(t) =
  \left( {\begin{array}{c}
   f_x(t)\\
   f_y(t)\\
  \end{array} } \right)
\end{equation}

where:

- $f_x(t)= \int_0^t u_x(t)\cdot dt$
- $f_y(t)= \int_0^t u_y(t)\cdot dt$ 
- $u_x(t)$ and $u_y(t)$ are the $\hat{i}$ and $\hat{j}$ components of point P's velocity

Similarly:

\begin{equation}
    Q(t) =
  \left( {\begin{array}{c}
   g_x(t) \\
      g_y(t) \\
  \end{array} } \right)
\end{equation}

where:

- $g_x(t)= \int_0^t v_x(t)\cdot dt$ 
- $g_y(t)= \int_0^t v_y(t)\cdot dt$
- $v_x(t)$ and $v_y(t)$ are the $\hat{i}$ and $\hat{j}$ components of point Q's velocity.

$P(t)$ and $Q(t)$ can be understood as the integrals of velocities $\vec{u}$ and $\vec{v}$ respectively over time t.

Additionally, we introduce adjusted TCA, a metric that measures the time P and Q avoided a collision by. To do so, we modify our calculation of P and Q to create $\hat{P}$ and $\hat{Q}$ that measure the delay in the behavior initiated at $t_0$ that would have led to a crash. Note, while the metric is perhaps most useful in circumstances where $t_0$ marks one vehicle's "reaction" to a threatening situation, it remains valid when no behavior change is initiated at $t_0$, which we refer to as a "null reaction".

\begin{equation}
    \hat{P}(t) =
  \left( {\begin{array}{c}
   f_x(t) + u_0\cdot\cos(\theta_0)\cdot Z_P\\
   f_y(t) + u_0\cdot\sin(\theta_0)\cdot Z_P\\
  \end{array} } \right)
\end{equation}

where: 

- $f_x(t)= \int_0^t u_x(t)\cdot dt$
- $f_y(t)= \int_0^t u_y(t)\cdot dt$
- $u_x(t)$ and $u_y(t)$ are the $\hat{i}$ and $\hat{j}$ components of point P's velocity
- $u_0$ is the velocity of point P at $t= 0$
- $\theta_0$ is the angle of rotation of P from $\hat{i}$ at $t=0$
- $Z_P$ is the delayed reaction time for point P, and

\begin{equation}
    \hat{Q}(t) =
  \left( {\begin{array}{c}
   g_x(t) - v_0\cdot\sin(\phi_0)\cdot Z_Q\\
   g_y(t) + v_0\cdot\cos(\phi_0)\cdot Z_Q\\
  \end{array} } \right)
\end{equation}

where:

- $g_x(t)= \int_0^t v_x(t)\cdot dt$ 
- $g_y(t)= \int_0^t v_y(t)\cdot dt$
- $v_x(t)$ and $v_y(t)$ are the $\hat{i}$ and $\hat{j}$ components of point Q's velocity
- $v_0$ is the velocity of point Q at $t= 0$
- $\phi_0$ is the angle of rotation of Q from $\hat{j}$ at $t=0$
- $Z_Q$ is the delayed reaction time for point Q

Since we assume that points $\hat{P}$ and $\hat{Q}$ crash at time t with delayed reactions $Z_P$ and $Z_Q$, we set $\hat{P}$ and $\hat{Q}$ equal to each other to get $Z_P$ and $Z_Q$ as functions of time. Here, time t denotes the time it takes for points P and Q to collide from their initial starting points.

### A Simple Example

The simplest example of our model assumes constant velocity over time:

\begin{equation}
    P(t) = P_0 + t\vec{u}
\end{equation} \
\begin{equation}
 Q(t) = Q_0 + t\vec{v}   
\end{equation}

where $P_0$ and $Q_0$ are positions of points P and Q at time $t = 0$.

Define a distance vector $\vec{d}(t)$  between P and Q as follows:

\begin{equation}
    \begin{split}
       \vec{d}(t) &= P(t) - Q(t) \\
       &= P_0 - Q_0 + (\vec{u} - \vec{v})t \\
       &= d_0 + (\vec{u} - \vec{v})t
    \end{split}
\end{equation}

where $d_0 = P_0 - Q_0$

We calculate the TCA of P and Q by minimizing $\vec{d}(t)$.

An animation of the scenario is as follows:
```{r, warning = FALSE, message = FALSE}
t <- 7
inc <- .01
p1 <- data.frame(x = rep(0,t/inc + 1), y = -50+10*seq(0,t, by = inc),  t = seq(0,t, by = inc))
q1 <- data.frame(x = 40 - 10*seq(0,t, by = inc), y = rep(0,t/inc + 1), t = seq(0,t, by = inc)) 
library(raster)
D <- pointDistance(p1[,-3], q1[,-3], lonlat = FALSE)
library(plotly)
df <- rbind(p1,q1)

plot1 <- df %>%
  plot_ly(
    x = ~x,
    y = ~y,
    frame = ~t,
    type = 'scatter',
    mode = 'markers',
    showlegend = F
  )
plot1
```

The TCA of the two points in this animation is:
```{r}
TCA1 <- which.min(D)
tca1 <- seq(0,t, by = inc)[TCA1]
tca1
```

Finally, the adjusted TCA gives us the graph:
```{r}
Zero <- data.frame(x = rep(0,t/inc + 1), y = rep(0,t/inc + 1))
DP <- pointDistance(p1[,-3], Zero, lonlat = FALSE)/rep(10, t/inc + 1)
DQ <- pointDistance(q1[,-3], Zero, lonlat = FALSE)/rep(-10, t/inc + 1)
plot(DP,DQ)
```

The area enclosed between (0, 1) and (-1, 0) is the measure of safety from collision.

**MORE INTERPRETATION/EXPLANATION OF THIS GRAPH IS NECESSARY, IT'S OUR MAIN RESULTS**

### Variable velocity with constant acceleration

We will modify our calculation of $\vec{u}$ as follows:

- Define the acceleration of P as $\vec{a_P} = a_{P0}\cos(\theta)\hat{i} + a_{P0}\sin(\theta)\hat{j}$

Adding acceleration to each component of velocity, we get:

\begin{equation}
    \vec{u}(t) = (u_0 \cos(\theta) + a_{P0}\cos(\theta)t)\hat{i} + (u_0\sin(\theta) + a_{P0}\sin(\theta)t)\hat{j}
\end{equation}

Similarly,

\begin{equation}
    \vec{v}(t) = -(v_0 \sin(\phi) + a_{Q0}\sin(\phi)t)\hat{i} + (v_0\cos(\phi) + a_{Q0}\cos(\phi)t)\hat{j}
\end{equation}

An animation of the scenario is as follows:
```{r}
p2 <- data.frame(x = rep(0,t/inc + 1), y = -25+10*seq(0,t, by = inc)-seq(0,t, by = inc)^2,  t = seq(0,t, by = inc))
q2 <- data.frame(x = 40 - 10*seq(0,t, by = inc), y = rep(0,t/inc + 1), t = seq(0,t, by = inc)) 
D <- pointDistance(p2[,-3], q2[,-3], lonlat = FALSE)
df <- rbind(p2,q2)
plot2 <- df %>%
  plot_ly(
    x = ~x,
    y = ~y,
    frame = ~t,
    type = 'scatter',
    mode = 'markers',
    showlegend = F
  )
plot2
```

The TCA of the two points in this animation is:
```{r}
TCA2 <- which.min(D)
tca2 <- seq(0,t, by = inc)[TCA2]
tca2
```

Finally, the adjusted TCA gives us the graph:
```{r}
Zero <- data.frame(x = rep(0,t/inc + 1), y = rep(0,t/inc + 1))
DP <- pointDistance(p2[,-3], Zero, lonlat = FALSE)/rep(10, t/inc + 1)
DQ <- pointDistance(q2[,-3], Zero, lonlat = FALSE)/rep(-10, t/inc + 1)
plot(DP,DQ)
```

### Non-zero angle of motion- fixed steering wheel angle

We use the steering wheel angle to determine the turn of wheels using the steering ratio. So, turn of wheels $=$ steering wheel angle $\times$ steering ratio.

Since turn of the wheel is the angle by which a vehicle turns at every instant, we interpret it as the derivative of a vehicle's angle of motion.

Let turn of wheel of point P be $W_P$. Then:

\begin{equation}
 W_P = \frac{d\theta}{dt}   
\end{equation}
Since we assume that $W_P$ is a constant, upon integrating both sides of equation (6) with respect to t, we get the angle of motion as a function of time as follows:
\begin{equation}
    \theta(t) = W_P \cdot t + \theta_0
\end{equation}

where $\theta_0$ is the angle at which P was turned at time $t = 0$.

Given a fixed steering wheel angle, point P travels with velocity $\vec{u}$ in a circular motion.

An animation of the scenario is as follows:
```{r}
p3 <- data.frame(x = -300/pi + 300/pi*cos(pi*seq(0,t, by = inc)/30), y = -150/pi+300/pi*sin(pi*seq(0,t,by=inc)/30),  t = seq(0,t, by = inc))
q3 <- data.frame(x = 40 - 10*seq(0,t, by = inc), y = rep(0,t/inc + 1), t = seq(0,t, by = inc)) 
D <- pointDistance(p3[,-3], q3[,-3], lonlat = FALSE)
df <- rbind(p3,q3)

plot3 <- df %>%
  plot_ly(
    x = ~x,
    y = ~y,
    frame = ~t,
    type = 'scatter',
    mode = 'markers',
    showlegend = F
  )
plot3
```

The TCA of the two points in this animation is:
```{r}
TCA3 <- which.min(D)
tca3 <- seq(0,t, by = inc)[TCA3]
tca3
```

Finally, the adjusted TCA gives us the graph:
```{r}
Zero <- data.frame(x = rep(-12.7936,t/inc + 1), y = rep(0,t/inc + 1))
DP <- pointDistance(p3[,-3], Zero, lonlat = FALSE)/rep(10, t/inc + 1)
DQ <- pointDistance(q3[,-3], Zero, lonlat = FALSE)/rep(-10, t/inc + 1)
plot(DP,DQ)
```

